Perturbing SU($n$)-symmetric interactions

### 2020-01-30 collective_sun/sun_perturbation.tex
commented out obsolete section on converting between multi-local and collective operators

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{multline}
  \P_0 \sp{\sum_{\substack{k<l\\p<q}}
    w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}} \P_0
  \EQFS \sum_{p<q} w_{Opq} w_{Qpq} O_{1,2} Q_{1,2}
  + \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}}
  w_{Okp} w_{Qkq} O_{1,2} Q_{1,3} \\
  + \sum_{\substack{k<\ell\\p<q\\\set{k,\ell}\ne\set{p,q}}}
  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
  \label{eq:POQP_start}
\end{multline}
where $O$ and $Q$ are permutationally-symmetric two-body operators;
and the coefficients $w_{Ok\ell},w_{Qpq}$ satisfy \eqref{eq:cond_2}
with eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
we can write the first sum on the right of \eqref{eq:POQP_start} as
\begin{align}
  \sum_{p<q} w_{Opq} w_{Qpq} = \v{\m w}_O \c\v{\m w}_Q,
\end{align}
and, using the fact that $w_{Opp}=0$, we can write the second sum on
the right of \eqref{eq:POQP_start} as
\begin{align}
  \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} w_{Okp} w_{Qkq}
  = \sum_{k,p} w_{Okp} \sum_{q\ne p} w_{Okq}
  = \sum_{k,p} w_{Okp} \p{w_{Qk} - w_{Qkp}}
  = \v w_O \c \v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Using additionally the fact that $w_{Opq},w_{Qpq}$ are mean-zero, we
can similarly simplify
\begin{align}
  \sum_{\substack{p<q\\k<\ell\\\set{k,\ell}\ne\set{p,q}}}
  w_{Ok\ell} w_{Qpq} = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{\substack{k<l\\p<q}} w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{1,3} \\
  - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{3,4}.
  \label{eq:POQP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$O_{1,2} Q_{2,3}$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \bar h_{pq} w_{pq}
  + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  &&
  \bar h_{pq} \equiv 2h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
which can be written in matrix form as
\begin{align}
  \bar{\m h} * \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
  = \Delta \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h$, $\bar{\m h}$, and $\m w$ are matrices of all $h_{pq}$,
$\bar h_{pq}$ and $w_{pq}$; $\bar{\m h}*\m w$ denotes an element-wise
(Kronecker) product of $\bar{\m h}$ and $\m w$, i.e.~with matrix
elements $\bar h_{pq}w_{pq}$; and
$\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
the projection of $\m m$ onto the space of strictly off-diagonal
matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
the interaction Hamiltonian $H_0$ is translationally invariant, in
which case $h_{pq}$ and $\bar h_{pq}$ depend only on the separation
$\pm\p{p-q}$, we expand all matrices in a basis that the respects
translational symmetry.  To this end, for all lattice displacements
$d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the
{\it cycle matrices}
\begin{align}
  \m m_{dk} \equiv \sum_p e^{ip\c k} \op{p+d}{p},
  \label{eq:cycle}
\end{align}
where the element-wise sums in $p+d$ are implicitly taken modulo $L$
(i.e.~the length of the periodic lattice along each axis).  In terms
of these cycle matrices, we define
\begin{align}
  \tilde w_{dk} \equiv \f1N \tr\p{\m m_{dk}^\dag \m w},
\end{align}
and expand
\begin{align}
  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
  &&
  \m h = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m m_{d,0},
  &&
  \bar{\m h} = \sum_d \bar h_{d,0} \m m_{d,0}.
\end{align}
We can then write the condition in \eqref{eq:cond_2_ref} as a new set
of conditions
\begin{align}
  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
  + \sum_{\substack{c,d\\c+d\ne0}}
  h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
\end{align}
that must hold for all wavenumbers $k$.  With some re-indexing, we can
write
\begin{align}
  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
  + \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
  \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
  \label{eq:cond_2_cycle}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}_k
  \equiv \sum_{d\ne0} \bar h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
  &&
  \ut{\v{\m w}}_k \equiv \sum_{d\ne0} \tilde w_{dk} \ket{d},
  \label{eq:cycle_vecs}
\end{align}
in order to express \eqref{eq:cond_2_cycle} as the set of eigenvalue
equations
\begin{align}
  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k = \Delta \ut{\v{\m w}}_k.
  \label{eq:cond_2_square}
\end{align}
The appearance of a wavenumber index $k$ in \eqref{eq:cond_2_square}
implies that all fixed-$k$ components of a general coefficient matrix
$\m w$ must simultaneously satisfy \eqref{eq:cond_2_square} in order
to generate states of fixed excitation energy $\Delta$ when the
corresponding perturbation $\V_2^\Delta$ acts on a fully symmetric
state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
coefficient vectors $\v{\m w}_\Delta$ that generate a fixed-energy
perturbation $\V_2^\Delta$ is then performed by projecting each
$k$-component of $\v{\m w}$ onto the corresponding $\Delta$-eigenspace
of $\ut{\check{\m h}}_k$.  If the perturbation $\V_2$ is
translationally invariant, then the $k$-components of $\v{\m w}$ for
all $k\ne 0$ all vanish, so we only need to consider the case of
$k=0$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Collective and multi-local operators in the fully symmetric
  manifold}
\label{sec:multi_to_collective}

Here we provide a decomposition multi-local operators into sums of
collective operators (and vice versa) within the fully symmetric
manifold of $N$ spins.  Multi-local operators are operators of the
form
\begin{align}
  O_S \equiv \bigotimes_{O\in S} O,
  \label{eq:multi_op_set}
\end{align}
where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
order of tensor factors in \eqref{eq:multi_op_set} does not matter
after projecting operators onto the fully symmetric manifold, as we
always will in this section.  While the decomposition of a collective
operator into a sum of multi-local operators is unique, the opposite
is not true: there is generally no unique decomposition of a
multi-local operator into a sum of collective operators.  We can,
however, enforce a canonical decomposition by choosing a set of
collective operators that obey the same symmetries as $O_S$, namely
invariance under arbitrary permutations of the single-spin operators
$O\in S$.

Even though $S$ contains $\abs{S}$ spins, a collective operator built
from the single-spin operators $O\in S$ may have multiple
(single-spin) operators acting on the same spin, resulting in a
$k$-local operator with $k<\abs{S}$.  We therefore {\it
  partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
  mutually disjoint, non-empty subsets $p\subset S$ with
  $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
$p\subset S$ is a subset of operators $O\in S$ that address the same
spin), we then define the $\abs{P}$-local operator
\begin{align}
  O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
\end{align}
where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
elements in $\X$, i.e.
\begin{align}
  \bigodot_{x\in\X} x
  \equiv \f1{\abs{\X}!}
  \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
  ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
\end{align}
with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
expressed in one-line notation; for example,
\begin{align}
  \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
  \bigodot_{j=1}^3 x_j
  &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
    + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
\end{align}
The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
consists of all operators that
\begin{enumerate*}
\item can be built from ordinary products and tensor products of all
  single-spin operators $O\in S$, and
\item are invariant under arbitrary permutations of $S$.
\end{enumerate*}
For any set $S$ of single-spin operators and any partition $P$ of $S$,
we can similarly define the collective operators
\begin{align}
  \O_S \equiv \bigodot_{O\in S} \col{O},
  &&
  \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
\end{align}
where we underline a single-spin operator to denote its collective
version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
operator that acts with the single-spin operator $Q$ on spin $j$ and
trivially on all other spins.  By construction, the set of collective
operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
and it is straightforward to work out that for any set $S$ of
single-spin operators,
\begin{align}
  \O_S
  \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
  N!_{\abs{P}} O_P,
  &&
  N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
  \label{eq:collective_to_multi_full}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold.  For any multi-local operator $O_S$ indexed by a
set $S$ of single-spin operators, we can collect expansions of the
form \eqref{eq:collective_to_multi_full} for all collective operators
built from non-empty subsets of $S$ into a system of linear equations
$\L$, and in turn solve $\L$ for $O_S$.  When
$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
a computer algebra system (namely, Mathematica) we find that
\begin{align}
  N!_{\abs{S}} O_S
  &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
  \label{eq:multi_to_collective_full} \\
  &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
  \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!

>>>>

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{multline}
  \P_0 \sp{\sum_{\substack{k<l\\p<q}}
    w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}} \P_0
  \EQFS \sum_{p<q} w_{Opq} w_{Qpq} O_{1,2} Q_{1,2}
  + \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}}
  w_{Okp} w_{Qkq} O_{1,2} Q_{1,3} \\
  + \sum_{\substack{k<\ell\\p<q\\\set{k,\ell}\ne\set{p,q}}}
  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
  \label{eq:POQP_start}
\end{multline}
where $O$ and $Q$ are permutationally-symmetric two-body operators;
and the coefficients $w_{Ok\ell},w_{Qpq}$ satisfy \eqref{eq:cond_2}
with eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
we can write the first sum on the right of \eqref{eq:POQP_start} as
\begin{align}
  \sum_{p<q} w_{Opq} w_{Qpq} = \v{\m w}_O \c\v{\m w}_Q,
\end{align}
and, using the fact that $w_{Opp}=0$, we can write the second sum on
the right of \eqref{eq:POQP_start} as
\begin{align}
  \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} w_{Okp} w_{Qkq}
  = \sum_{k,p} w_{Okp} \sum_{q\ne p} w_{Okq}
  = \sum_{k,p} w_{Okp} \p{w_{Qk} - w_{Qkp}}
  = \v w_O \c \v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Using additionally the fact that $w_{Opq},w_{Qpq}$ are mean-zero, we
can similarly simplify
\begin{align}
  \sum_{\substack{p<q\\k<\ell\\\set{k,\ell}\ne\set{p,q}}}
  w_{Ok\ell} w_{Qpq} = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{\substack{k<l\\p<q}} w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{1,3} \\
  - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{3,4}.
  \label{eq:POQP}
\end{multline}

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \bar h_{pq} w_{pq}
  + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  &&
  \bar h_{pq} \equiv 2h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
which can be written in matrix form as
\begin{align}
  \bar{\m h} * \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
  = \Delta \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h$, $\bar{\m h}$, and $\m w$ are matrices of all $h_{pq}$,
$\bar h_{pq}$ and $w_{pq}$; $\bar{\m h}*\m w$ denotes an element-wise
(Kronecker) product of $\bar{\m h}$ and $\m w$, i.e.~with matrix
elements $\bar h_{pq}w_{pq}$; and
$\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
the projection of $\m m$ onto the space of strictly off-diagonal
matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
the interaction Hamiltonian $H_0$ is translationally invariant, in
which case $h_{pq}$ and $\bar h_{pq}$ depend only on the separation
$\pm\p{p-q}$, we expand all matrices in a basis that the respects
translational symmetry.  To this end, for all lattice displacements
$d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the
{\it cycle matrices}
\begin{align}
  \m m_{dk} \equiv \sum_p e^{ip\c k} \op{p+d}{p},
  \label{eq:cycle}
\end{align}
where the element-wise sums in $p+d$ are implicitly taken modulo $L$
(i.e.~the length of the periodic lattice along each axis).  In terms
of these cycle matrices, we define
\begin{align}
  \tilde w_{dk} \equiv \f1N \tr\p{\m m_{dk}^\dag \m w},
\end{align}
and expand
\begin{align}
  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
  &&
  \m h = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m m_{d,0},
  &&
  \bar{\m h} = \sum_d \bar h_{d,0} \m m_{d,0}.
\end{align}
We can then write the condition in \eqref{eq:cond_2_ref} as a new set
of conditions
\begin{align}
  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
  + \sum_{\substack{c,d\\c+d\ne0}}
  h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
\end{align}
that must hold for all wavenumbers $k$.  With some re-indexing, we can
write
\begin{align}
  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
  + \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
  \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
  \label{eq:cond_2_cycle}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}_k
  \equiv \sum_{d\ne0} \bar h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
  &&
  \ut{\v{\m w}}_k \equiv \sum_{d\ne0} \tilde w_{dk} \ket{d},
  \label{eq:cycle_vecs}
\end{align}
in order to express \eqref{eq:cond_2_cycle} as the set of eigenvalue
equations
\begin{align}
  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k = \Delta \ut{\v{\m w}}_k.
  \label{eq:cond_2_square}
\end{align}
The appearance of a wavenumber index $k$ in \eqref{eq:cond_2_square}
implies that all fixed-$k$ components of a general coefficient matrix
$\m w$ must simultaneously satisfy \eqref{eq:cond_2_square} in order
to generate states of fixed excitation energy $\Delta$ when the
corresponding perturbation $\V_2^\Delta$ acts on a fully symmetric
state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
coefficient vectors $\v{\m w}_\Delta$ that generate a
fixed-excitation-energy perturbation $\V_2^\Delta$ is then performed
by projecting each $k$-component of $\v{\m w}$ onto the corresponding
$\Delta$-eigenspace of $\ut{\check{\m h}}_k$.  If the perturbation
$\V_2$ is translationally invariant, then the $k$-components of
$\v{\m w}$ for all $k\ne 0$ all vanish, so we only need to consider
the case of $k=0$.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Collective and multi-local operators in the fully symmetric
%   manifold}
% \label{sec:multi_to_collective}

% Here we provide a decomposition multi-local operators into sums of
% collective operators (and vice versa) within the fully symmetric
% manifold of $N$ spins.  Multi-local operators are operators of the
% form
% \begin{align}
%   O_S \equiv \bigotimes_{O\in S} O,
%   \label{eq:multi_op_set}
% \end{align}
% where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
% order of tensor factors in \eqref{eq:multi_op_set} does not matter
% after projecting operators onto the fully symmetric manifold, as we
% always will in this section.  While the decomposition of a collective
% operator into a sum of multi-local operators is unique, the opposite
% is not true: there is generally no unique decomposition of a
% multi-local operator into a sum of collective operators.  We can,
% however, enforce a canonical decomposition by choosing a set of
% collective operators that obey the same symmetries as $O_S$, namely
% invariance under arbitrary permutations of the single-spin operators
% $O\in S$.

% Even though $S$ contains $\abs{S}$ spins, a collective operator built
% from the single-spin operators $O\in S$ may have multiple
% (single-spin) operators acting on the same spin, resulting in a
% $k$-local operator with $k<\abs{S}$.  We therefore {\it
%   partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
%   mutually disjoint, non-empty subsets $p\subset S$ with
%   $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
% the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
% $p\subset S$ is a subset of operators $O\in S$ that address the same
% spin), we then define the $\abs{P}$-local operator
% \begin{align}
%   O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
% \end{align}
% where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
% elements in $\X$, i.e.
% \begin{align}
%   \bigodot_{x\in\X} x
%   \equiv \f1{\abs{\X}!}
%   \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
%   ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
% \end{align}
% with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
% expressed in one-line notation; for example,
% \begin{align}
%   \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
%   \bigodot_{j=1}^3 x_j
%   &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
%     + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
% \end{align}
% The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
% consists of all operators that
% \begin{enumerate*}
% \item can be built from ordinary products and tensor products of all
%   single-spin operators $O\in S$, and
% \item are invariant under arbitrary permutations of $S$.
% \end{enumerate*}
% For any set $S$ of single-spin operators and any partition $P$ of $S$,
% we can similarly define the collective operators
% \begin{align}
%   \O_S \equiv \bigodot_{O\in S} \col{O},
%   &&
%   \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
% \end{align}
% where we underline a single-spin operator to denote its collective
% version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
% operator that acts with the single-spin operator $Q$ on spin $j$ and
% trivially on all other spins.  By construction, the set of collective
% operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
% set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
% and it is straightforward to work out that for any set $S$ of
% single-spin operators,
% \begin{align}
%   \O_S
%   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
%   N!_{\abs{P}} O_P,
%   &&
%   N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
%   \label{eq:collective_to_multi_full}
% \end{align}
% where $\EQFS$ denotes equality under a restriction to the fully
% symmetric manifold.  For any multi-local operator $O_S$ indexed by a
% set $S$ of single-spin operators, we can collect expansions of the
% form \eqref{eq:collective_to_multi_full} for all collective operators
% built from non-empty subsets of $S$ into a system of linear equations
% $\L$, and in turn solve $\L$ for $O_S$.  When
% $\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
% a computer algebra system (namely, Mathematica) we find that
% \begin{align}
%   N!_{\abs{S}} O_S
%   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
%   \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
%   \label{eq:multi_to_collective_full} \\
%   &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
